---
title: "Motor Trend Analysis"
author: "Gulab Singh"
date: "27 March 2017"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Executive Summary
In this report we will analyse the dataset for a collection of manual and automatic transmission and study the relationship between a set of variables and fuel effeciency expressed in Miles per Gallon (MPG). The study is focussed in finding answers to following questions

* Is an automatic or manual transmission better for MPG
* Quantify the MPG difference between automatic and manual transmissions

## Exploratory data analysis

```{r exploratory}
library(datasets)
data("mtcars")
str(mtcars[mtcars$am=="0",])
str(mtcars[mtcars$am=="1",])
```

There are a total of 32 observatiosn fo 11 variables, 19 observations are for automatic transmission (am=0) and 13 for mannual transmission (am=1), looking at the summary data, the mean of mpg for mannual cars is 7.24 more than for automatic cars (24.39-17.15). This is also illustrated in the boxplot included in Appendix A, however we will need to carry out hypothesis testing and also look at the impact of variables other than transmission type on mpg.

## Regression analysis
To bettwer understand the imapct of various variables, we will plot a relationship between all the variables (refer Appendix A). 

Explortaory analysis suggested that the mpg for manual cars is higher as compared to automatic transmission. we will test this using t-test.

Null Hypothsis: There is no difference in MPG between Automatic and Manual transmission cars
Alternate Hypothsis: There is difference in MPG between Automatic and Manual transmission cars

```{r nullhypo}
t.test(mtcars$mpg~mtcars$am, conf.level=0.95)

```

Since p-value is 0.001374, we ust reject the Null Hypothesis

```{r lm}
linearfit <- lm(formula = mpg ~ am, data = mtcars)
summary(linearfit)
```

From above we can conclude that the fuel effecienncy for automatic cars is 17.147 and manual cars run at 17.147 + 7.245 = 24.392

since R squared value is 0.3598, this accounts for only 36% variance. We will now perform a multi variate linera regression analysis

We will let the model choose most significant paramaters.

```{r multi}
automultivariate <- step(lm(data=mtcars, mpg ~ .), trace=0)
summary(automultivariate)

```

Considering wt, qsec and am explains 85% of the variance. Residual plots are included in Appendix A

## Results

Car Weight and Acceleration must be considered to study impact of transmission type (Automatic or Manual) on the fuel effeciency (MPG) of the car

## Appendix A

```{r plot1}
boxplot(mpg ~ am, data = mtcars, xlab = "Transmission Type", ylab = "Fuel Effeciency (mpg)", col="gray",border = c("blue","green"), main="Fuel Effeciency by Transmission Type") 

pairs(mpg ~ ., data = mtcars, main="plot 2 - Exploring relationships between all pairs of variables")

plot(automultivariate)
```

